Handling String-Number Conversion in String Constraint Solving

### 2019-10-18 draft.tex
add disequality

String data type is omnipresent in modern programming languages. Various infamous security vulnerabilities such as injection and cross-site scripting attack are caused by malicious string values. As a results, in the past, a significant amount of research efforts have been investigated to symbolic execution and test case generation of string manipulating programs~\cite{saxena2010symbolic,artzi2011framework,huang2004securing,sen2013jalangi}. The core enabling technology for these approaches is string constraint solving~\cite{kiezun2009hampi,zheng2013z3,abdulla2015norn,abdulla2017flatten,wang2016string,abdulla2018trau,chen2019decision,zheng2017z3str2}. However, all these modern solvers fall short when \textit{string-number conversion function} occurs in the program.

@@

\section{Handling Equality and Disequality Constraint} \label{section:eq}
Given an equality constraint $x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$, we restrict the domain of all variables to some SFA with disjoint alphabet and then connect the final state and initial state of two consecutive variables with an epsilon transition. We rename the variables on the transitions of SFA's for different occurrences of the same variable in one equality constraint. Later we will use some arithmetic constraints to enforce that all occurrences of the variants of the same variable will be assigned the same value. For notational convenience, the renaming is done by adding a version number to the variable. For example, we can rename $v_1$ to $v_1^{(0)}$, $v_1^{(1)}$, $\ldots$.

@@

>>>>

String data type is omnipresent in modern programming languages. Various infamous security vulnerabilities such as injection and cross-site scripting attack are caused by malicious string values. As a results, in the past, a significant amount of research efforts have been investigated to symbolic execution and test case generation of string manipulating programs~\cite{saxena2010symbolic,artzi2011framework,huang2004securing,sen2013jalangi}. The core enabling technology for these approaches is string constraint solving~\cite{kiezun2009hampi,abdulla2014string,zheng2013z3,abdulla2015norn,abdulla2017flatten,wang2016string,abdulla2018trau,chen2019decision,zheng2017z3str2}. However, all these modern solvers fall short when \textit{string-number conversion function} occurs in the program.

@@

\section{Handling Equality and Disequality Constraint Symbolically} \label{section:eq}

Given an equality constraint $x_1\cdot x_2 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$, we restrict the domain of all variables to some SFA with disjoint alphabet and then connect the final state and initial state of two consecutive variables with an epsilon transition. We rename the variables on the transitions of SFA's for different occurrences of the same variable in one equality constraint. Later we will use some arithmetic constraints to enforce that all occurrences of the variants of the same variable will be assigned the same value. For notational convenience, the renaming is done by adding a version number to the variable. For example, we can rename $v_1$ to $v_1^{(0)}$, $v_1^{(1)}$, $\ldots$.

@@

Disquality constraint $x_1\cdot x_2 \cdots x_n \neq x_{n+1}\cdot x_{n+2} \cdots x_m$ can be converted to equal-satisfiable equality constraints in the standard way~\cite{abdulla2015norn}. We replace it with the following constraints.

$$
\begin{array}{c}
(|x_1|+ |x_2|+ \cdots +|x_n| \neq |x_{n+1}| + |x_{n+2}|+ \cdots +|x_m|)\vee \\
\left(
	\begin{array}{cccc}

	 y_1\cdot y_2\cdot y_3 &=& x_1\cdot x_2 \cdots x_n &\wedge\\
	 y_1 \cdot y'_2 \cdot y'_3 &=& x_{n+1}\cdot x_{n+2} \cdots x_m &\wedge\\
	|y_2|&=&|y'_2|=1
	\end{array}

\right)
\end{array}
$$

The second disjunct says that the constraints $x_1\cdot x_2 \cdots x_n$ and $x_{n+1}\cdot x_{n+2} \cdots x_m$ have a common prefix $y_1$, but the next character $y_2$ and $y'_2$ is different. In an more efficient implementation, one just project $y_2$ and $y'_2$ to a SFA that accept only symbolic words $v_1$ and $v'_1$, respectively (the SFA has no loop and only one transition).
